\documentclass[12pt]{article}

\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{changepage}
\usepackage{graphicx}
  
\title{Worksheet 2: SMP Reductions}
\author{Rin Meng 51940633 \\ Kevin Zhang 10811057 \\ Mika Panagsagan 29679552 \\ Priyansh Mathur 84491356}
\date{\today}

\begin{document}

\maketitle

\setcounter{section}{5}

\section{Promising Approach}

A group of residents each needs a residency in some hospital. A group of hospitals each need some
number (one or more) of residents, with some hospitals needing more and some fewer. Each group has
preferences over which member of the other group they’d like to end up with. The total number of slots in
hospitals is exactly equal to the total number of residents.
We want to fill the hospitals slots with residents in such a way that no resident and hospital that
weren’t matched up will collude to get around our suggestion (and give the resident a position at that
hospital instead).

\begin{enumerate}
    \item We can reduce problem $B$ into SMP problem.
    \item We can set up preference list for each hospitals and residents. For example we can have $n$ residents
    and hospitals can have $n$ or $n - 1$.
        \begin{itemize}
            \item Resident's $R_n$ preference list can have up to $n$ or $n-1$ number of hopsitals.
            \item Hospital's, $H_n$ prefference list can have up to $n$ residents.
            \item $R_1$: $H_1, H_2$
            \item $R_2$: $H_2, H_1$
            \item $R_3$: $H_1, H_2$
            \item $H_1$: $R_1, R_2, R_3$
            \item $H_2$: $R_2, R_1, R_3$
        \end{itemize}
    \item The solution to the RHP instance from our solution $B$ instance, we can make a 
     stable matching set is $M = (R_1, H_1), (R_2, H_2), (R_3, H_1)$
    \item To convert from any instance $I$ of RHP into an instance $I'$ of $B$,
    \begin{enumerate}
        \item We convert this problem to an SMP problem.
        \item For each resident and hospitals, we give them a preference list.
        \item Then we make an empty stable matching list $M$.
        \item We follow GS algorithm.
        \item We can convert the stable matching list $M$ to a solution to the RHP instance.
    \end{enumerate}
    \item To convert from any instance $I$ of $B$ into an instance $I'$ of RHP,
    \begin{enumerate}
        \item Since our RHP set is still empty, we can implement the GS algorithm.
        \item Residents has its own preference list as well as hospitals.
        \item Find the stable match for either residents or hospitals.
        \item We iterate until we get the most stable match.
    \end{enumerate}
\end{enumerate}


\end{document}
